---
title: "A4; incarceration rate"
author: "Samuel shumye"
date: '2022-11-01'
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
The American legal system is severely corrupt and biased. Along with police violence, the unequal incarceration rate is a daily occurrence of injustice. In this assignment, I choose to analyze incarceration statistics to demonstrate the disparity in incarceration rates between racial groups in the United States. I began by gaining a general understanding of the incarceration rate across the years. To examine the unjust system, I had to visualize the incarceration rates of different races, in this case, black and white. I will conclude my data analysis by showing the incarceration map of a single state.

## Introduction 
The issue with the united states prison system is that it is morally ineffective, as well as racially biased. In fact, black Americans are incarcerated at a state average of 1,240 per 100,000 residents, whereas Latino Americans are imprisoned at a rate of 349 per 100,000 residents. White, non-Latino Americans, meanwhile, are incarcerated at 261 per 100,000 residents. To add, 76 percent of all inmates end up back in jail within five years. In order to solve the problem, we need to identify the issue through data. This is what I will be trying to go through this assignment. This data provides useful information for visualizing and understanding how to address the difficulties we confront in modern society. People of different races have a direct stake in this information, as do corporations that utilize criminal histories to hire personnel. Indirect stakeholders who benefit from this data may include courts and justice departments.



## Data Summary 
In the summary, I decided it was a good Idea to find information that would prove to be somewhat impactfull. These include of creating a summary of black people to white people in jail, the total, average, max, and min. In addition to that information, I included information on a summary for the maximum black and white people in jail in the state of alabama, as this is where my family resides. I feel that all the information I collected would give the viewer an idea on the number bias due to lower population rate, yet higher jail population.
  


## Growth of the U.S. Prison Population 

```{r, echo=FALSE, warning=FALSE, message=TRUE, error=TRUE}
 results
```

In the past 50 years, the jail population has nearly quadrupled, increasing by more than 400%. As shown by the graph beginning in 1970, the number of convicts has gradually increased at a rapid rate. What factors contribute to the rate of incarceration rapid increase? What role do congressional crimes acts play in this circumstance?
## Growth of Prison Population by State

```{r, echo=FALSE, warning=FALSE, message=TRUE, error=TRUE}
task
```
We can see a noticable change on the states from the chart. California has the highest incarceration rate, followed by Washington and Oregon. What is the reason for these differences we see on the chart and does state laws affect the outcomes of U.S prison syatem?
## inequality chart

```{r, echo=FALSE, warning=FALSE, message=TRUE, error=TRUE}
determine
```
Even though 13.6% of the U.S. population is black, they make up 38% of the prison population. The data shown above demonstrates that the black incarceration rate is significantly greater than that of whites. What factors contribute to the jail system's racial disparity? Can we do anything to improve the situation?

## map

```{r, echo=FALSE, warning=FALSE, message=TRUE, error=TRUE}
 cases_Map
```



Texas has one of the highest incarceration rates in the nation. In Texas, there are approximately 135 000 prisoners. By dividing the prisoner population by county, it is easier to determine which area has the highest incarceration rate and to inquire whether this is a result of a biased prison system. Based on this map, do we have sufficient evidence to conclude that the system is unequal?
_All the very best with your work!_



